Publilia-03-survivorship.Rmd

### 2023-08-11 scripts/Publilia-03-survivorship.Rmd
Moved all data prep for `Publilia-03-survivorship.Rmd` to `Publilia-01-qc.Rmd`

# 0 - Data import

clutch_status_guarding_summary_v2
```{r}
All_data_v6 <- read_rds(here("saved_rds/All_data_v6.rds"))

clutch_status_guarding_summary_v2 <- read_rds(here("saved_rds/clutch_status_guarding_summary_v2.rds"))
# last_day_fID <- read_rds(here("saved_rds/last_day_fID.rds"))
# 
# strats_data <- left_join(clutch_status_guarding_summary_v2, last_day_fID)
```

# Final figures and stats using functions
Use the same data for each regression
```{r}
# I originally used the ants_removed as the predictor representing ants, but I am now redoing the analysis using whether ants were ever observed at the clutch.
surv_reg_data <- clutch_survivorship_data %>% 
  filter(!is.na(`visited in 4 days`)) %>%
  select(location, med_t5eggs, n_hatched, n_survived,
         des=last_des_rel2h, ants_removed, vis=`visited in 4 days`,
         females_removed, initiation_date, est_hatch) %>%
  mutate(n_survived = ifelse(n_survived/n_hatched > 1,# if n_survived/n_hatched>1
                             n_hatched, n_survived), # force n_survived/n_hatched=1
         ant_treatment=ifelse(ants_removed, "Removed", "Control")) %>%
  left_join(select(ant_counts, location, 
                   ever_antsP_b4h, ever_antsP_afh, ever_antsP)) %>%
  mutate(ants_b4h=ifelse(ever_antsP_b4h," ever", " never"),
         ants_afh=ifelse(ever_antsP_afh," ever"," never"),
         ants_ever=ifelse(ever_antsP," ever"," never")) %>%
  left_join(select(ant_n_days, -ants_everP, -ant_treatment, -females_removed)) %>%
  mutate(ants_gt2d=ifelse(antsP_gt2d," ants", " no ants"),
         ants_gt3d=ifelse(antsP_gt3d," ants", " no ants")) %>% 
  separate(location, into = c("DormPlant", "Leaf"), sep="_", remove=FALSE) %>%
  left_join(dplyr::select(moms_plant_data, DormPlant, last_visit_ends_P), by="DormPlant") %>%
  mutate(last_desP_rel2h = last_visit_ends_P-est_hatch,
         desP = last_desP_rel2h) 
```

# Final (really this time) figures for the dissertation
Choosing the models strictly based on qAICc leads to some odd results that make biological interpretation too complciated.

Hatching success: All model types are within ~2 qAICc of each other, so pick whichever shows the biology best and is most consistent with Nsurv and Esurv.

Nymph survivorship: desP * ants_ever * desP^2 has the best qAICc, and the desL version has the second best. The desL version (which, for biological interpretation's sake, should be in the main figure) has very wide 95% CI on the left side, which sends off alarm bells that model fit may be a less robust to addition or subtraction of samples than the AIC suggests.

Egg survivorship: desL * ants_ever * desL^2  has best qAICc, by far. Again the left side 95% CI is huge and it just looks terrible, suggesting to me that there may be too few influential points on the right side to conclusively support the convex up shape.

One problem with the ants_ever method is that the `never ants` level has a much smaller sample size than the `ever ants` level. I'm worried that this could make conclusive inference on the difference among these treatments or even just slope and shape could be shaky. Breaking the groups into `ants >2 days` vs `ants <=2 days` solves the unequal sample sizes problem and makes the desL figure more straightforward to interpret. The qAICc is considerably worse, but I'm concerned that it is only because ants ever was overfitting with too little data.

So, I'm making `desL * ants_gt2 * desL^2` the main figure and `desP * ants_gt2 * desP^2` the supplemental figure. I will report the qAICc of these two models, but not in a table. The data summary table, regression table and results text will need to change too.

*******

After running all these new regressions, I realize that the desL * ant treatment quadratic model has the lowest qAICc and the most interpretable plot and summary stats. I will use this instead.

And since the desP models perform poorly (both in terms of AICc and a convex up Nsurv fit), I just won't even include them in the supplement. I can include the desL `ants >2 days` in the supplement, but honestly it might not even be necessary since it looks so similar to the treatment-based model. I can just mention they are similar?

des_hatch.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data_LastDesPlant, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desL.ants_gt2d.quad.fits$mai.qb)

des_Nsurv.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data_LastDesPlant, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desL.ants_gt2d.quad.fits$mai.qb)

des_Esurv.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data_LastDesPlant, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_survived, survDen=med_t5eggs,
              model = des_Esurv.desL.ants_gt2d.quad.fits$mai.qb)

@@

des_hatch.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data_LastDesPlant, 
              tenure=des, treatment=ant_treatment,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desL.ant_treatment.quad.fits$mai.qb)

@@

des_Nsurv.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data_LastDesPlant, 
              tenure=des, treatment=ant_treatment,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desL.ant_treatment.quad.fits$mai.qb)

@@

des_Esurv.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data_LastDesPlant, 
              tenure=des, treatment=ant_treatment,
              survNom=n_survived, survDen=med_t5eggs,
              model = des_Esurv.desL.ant_treatment.quad.fits$mai.qb)

@@

@@

des_hatch.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data_LastDesPlant, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desL.ants_gt2d.quad.fits$mai.qb)

@@

des_Nsurv.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data_LastDesPlant, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desL.ants_gt2d.quad.fits$mai.qb)

>>>>

# 0 - Data import and custom functions

## Data for regressions
```{r}
surv_reg_data <- read_rds(here("saved_rds/surv_reg_data.rds"))
```

## Custom functions

### Model fitting and confidence intervals and plotting functions
```{r}
### fit survivorship model
fit_surv <- function(data_in, survNom, survDen, tenure, 
                     treatment, vis, interaction=c("full","pw")){
  data <- eval(substitute(data_in %>% filter(!is.na(tenure))))

  mai.b <- eval(substitute(glm(survNom/survDen ~ 
            tenure*treatment,
          data = data, weights = survDen, family = binomial)))
  print("Interaction, Binomial: ")
  print(capture.output(summary(mai.b))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(mai.b))))-1)])

  mai.qb <- eval(substitute(glm(survNom/survDen ~ 
            tenure*treatment,
          data = data, weights = survDen, family = quasibinomial)))
  print("Interaction, Quasibinomial: ")
  print(capture.output(summary(mai.qb))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(mai.qb))))-1)])

  ma.qb <- eval(substitute(glm(survNom/survDen ~ 
            tenure+treatment,
          data = data, weights = survDen, family = quasibinomial)))
  print("No Interaction, Quasibinomial: ")
  print(capture.output(summary(mai.qb))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(mai.qb))))-1)])

  mai.models <- list(mai.b = mai.b, mai.qb = mai.qb, ma.qb = ma.qb)
  return(mai.models)
}

# Example
# test_fits <- fit_surv(data_in=hatch.data, tenure=des, treatment=ants,
#                  interaction="pw", survNom=n_hatched, survDen=med_t5eggs)

### fit survivorship model
fit_surv_quad <- function(data_in, survNom, survDen, tenure, 
                     treatment){
  data <- eval(substitute(data_in %>% filter(!is.na(tenure))))

  mai.b <- eval(substitute(glm(survNom/survDen ~
            tenure*treatment + (I(as.numeric(tenure)^2))*treatment,
          data = data, weights = survDen, family = binomial)))
  print("Interaction, Binomial: ")
  print(capture.output(summary(mai.b))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(mai.b))))-1)])

  mai.qb <- eval(substitute(glm(survNom/survDen ~
            tenure*treatment + (I(as.numeric(tenure)^2))*treatment,
          data = data, weights = survDen, family = quasibinomial)))
  print("Interaction, Quasibinomial: ")
  print(capture.output(summary(mai.qb))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(mai.qb))))-1)])

  ma.qb <- eval(substitute(glm(survNom/survDen ~
            tenure+treatment + (I(as.numeric(tenure)^2))*treatment,
          data = data, weights = survDen, family = quasibinomial)))
  print("No Interaction, Quasibinomial: ")
  print(capture.output(summary(ma.qb))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(ma.qb))))-1)])

    mai.models <- list(mai.b = mai.b, mai.qb = mai.qb, ma.qb = ma.qb)
  return(mai.models)
}

# Example
# test_fits <- fit_surv(data_in=hatch.data, tenure=des, treatment=ants,
#                  survNom=n_hatched, survDen=med_t5eggs)

### Create confidence intervals in the link scale for plotting, and backtransform for plot
# Based on: https://fromthebottomoftheheap.net/2018/12/10/confidence-intervals-for-glms/
ci_for_plot <- function(data_in, survNom, survDen, 
                        tenure, treatment, model){

  data <- eval(substitute(
    data_in %>% filter(!is.na(tenure)) %>%
      select(location, tenure, treatment, survDen, survNom) %>%
      add_column(fit = predict(model, newdata = ., type = 'response'))
    ))

# Example
# test_fits.pw <- fit_surv(data_in=hatch.data, tenure=des, treatment=ants, 
#                  interaction="pw", survNom=n_hatched, survDen=med_t5eggs)
# 
# test_ci_data <- ci_for_plot(data_in=hatch.data, tenure=des, treatment=ants, 
#                  survNom=n_hatched, survDen=med_t5eggs,
#                  model = test_fits.pw$mai.qb)

### Plot model results with confidence intervals
plot_ci <- function(ci_data, survNom, survDen, treatment, tenure){
  (eval(substitute(plot <- ggplot(data = ci_data, 
       aes(x = tenure, y = fit, 
           color = treatment, fill = treatment)) +
    theme_publilia() +
    geom_point(aes(y = survNom/survDen), alpha = 0.25, size = 4, shape=16, key_glyph=draw_key_rect) + 
    geom_line(aes(y = fit), size = 1, key_glyph=draw_key_rect) +
    geom_ribbon(aes(ymin = conf_upr, ymax = conf_lwr),
                    alpha = 0.5, color = NA, key_glyph=draw_key_rect))))
#
}

# Example
# test_fits.pw <- fit_surv(data_in=hatch.data, tenure=des, treatment=ants,
#                  interaction="pw", survNom=n_hatched, survDen=med_t5eggs)
# 
# test_ci_data <- ci_for_plot(data_in=hatch.data, tenure=des, treatment=ants,
#                  survNom=n_hatched, survDen=med_t5eggs,
#                  model = test_fits.pw$mai.qb)
# 
# (p <- plot_ci(ci_data = test_ci_data, tenure = des, treatment=ants,
#              survNom = n_hatched, survDen = med_t5eggs) +
#     xlab("Latest desertion date relative to hatch") +
#     ylab("Hatching success") +
#     scale_color_manual(values = c("brown1", "black")) +
#     scale_fill_manual(values = c("brown1", "black")))
# summary(test_fits.pw$mai.qb)
```

@@

des_hatch.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ant_treatment,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desL.ant_treatment.quad.fits$mai.qb)

@@

des_Nsurv.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ant_treatment,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desL.ant_treatment.quad.fits$mai.qb)

@@

des_Esurv.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ant_treatment,
              survNom=n_survived, survDen=med_t5eggs,
              model = des_Esurv.desL.ant_treatment.quad.fits$mai.qb)

@@

des_hatch.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desL.ants_gt2d.quad.fits$mai.qb)

des_Nsurv.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desL.ants_gt2d.quad.fits$mai.qb)

des_Esurv.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_survived, survDen=med_t5eggs,
              model = des_Esurv.desL.ants_gt2d.quad.fits$mai.qb)

@@

des_hatch.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desL.ants_gt2d.quad.fits$mai.qb)

@@

des_Nsurv.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desL.ants_gt2d.quad.fits$mai.qb)
